### 2020-08-14 TCS_v3/AnswerToReviewer.txt

Dear Editor and Referees,

we thank you very much for your detailed comments and remarks, which greatly helped us to correct and improve our presentation. 
We updated the manuscript following the advices of the referees. Please find below a detailed description of the way in which 
the items pointed out by the referees were addressed in the revised version.

Best regards,
Linda Brodo, Roberto Bruni, Moreno Falaschi

-------------------------

Manuscript Number: TCS-D-20-00267R1
Manuscript Title: A process algebraic approach to reaction systems.
Author(s): Linda Brodo, Researcher; Roberto Bruni, Professor; Moreno Falaschi, Professor
Submitted to: Theoretical Computer Science

-------------------------
Reviewer #1: I think my original comments have been addressed satisfactorily by the authors, but I still have some remarks:

p.6, l.12 - replace "set of entities" by "sets of entities"
> OK. Done

p.7, l.42 - replace "where v_i is a link chain" by "where v is a link chain" 
as in the syntax from the previous line you use v.P and not v_i.P
> OK. Done

p.8, l.13, (Res) rule - make sure all a's have same font as it seems the one in (\nu a)P' is different
> OK. Done

p.8, l.30 - replace "Fig.1" by "Fig. 1"
> OK. Done

p.12, l.35 (item 2 of Lemma 12) - replace "for each reaction" by "each reaction"
> OK. We have corrected the formulation.

p.12, l.51 - in Def. 13 shouldn't be P \xrightarrow{(\nu names)v} P' instead of P \xrightarrow{v}P' ?  Only starting on p.15, l.52 you say "In this section, for simplicity, we shall often omit topmost restrictions (\nu names)". Thus, either move this assumption earlier or till Section 5 keep using (\nu names) on \xrightarrow

p.13, l.56 - (\nu names) missing from \xrightarrow label

p.17, l.49 - replace "(starting at cxt and ending at p1" by "(starting at cxt and ending at p1)", namely add missing bracket

p.33, l.41 - shouldn't be "j \geq 1" (as assumed on p.33, l.38) instead of "j \geq 0"?

p.33, l.47 and l.49 - in the proof of the base case "j=1", instead of "j-1" shouldn't be "j"? This is due to the fact that Prop. 16 you mention you have the case j=1 and also on p.33 on l.52 you got \gamma^j and not \gamma^{j-1}.

### 2020-08-16 TCS_v3/AnswerToReviewer.txt

p.8, l.13, (Res) rule - make sure all a's have same font as it seems the one in (\nu a)P' is different
> OK. Done

p.8, l.30 - replace "Fig.1" by "Fig. 1"
> OK. Done

p.12, l.35 (item 2 of Lemma 12) - replace "for each reaction" by "each reaction"
> OK. We have corrected the formulation.

p.12, l.51 - in Def. 13 shouldn't be P \xrightarrow{(\nu names)v} P' instead of P \xrightarrow{v}P' ?  Only starting on p.15, l.52 you say "In this section, for simplicity, we shall often omit topmost restrictions (\nu names)". Thus, either move this assumption earlier or till Section 5 keep using (\nu names) on \xrightarrow
>  ***TO BE ANSWERED***

p.13, l.56 - (\nu names) missing from \xrightarrow label
>  ***TO BE ANSWERED***

p.17, l.49 - replace "(starting at cxt and ending at p1" by "(starting at cxt and ending at p1)", namely add missing bracket
> OK. Done

p.33, l.41 - shouldn't be "j \geq 1" (as assumed on p.33, l.38) instead of "j \geq 0"?
> OK. Done

p.33, l.47 and l.49 - in the proof of the base case "j=1", instead of "j-1" shouldn't be "j"? This is due to the fact that Prop. 16 you mention you have the case j=1 and also on p.33 on l.52 you got \gamma^j and not \gamma^{j-1}.
>  ***TO BE DONE/ANSWERED***

### 2020-08-16 TCS_v3/06-examples.tex
cambi vari

Then, only the second reaction can be applied, and the transition carries the complete label $\upsilon$ below
{\tiny
\[
%\begin{array}{l}
%\upsilon = \\
%\restrict{r1,r2,r3,r4,\widetilde{q},\widetilde{w},\widetilde{a},\widetilde{b}}\\
 \startchain{\silent}
 \chainedlink{r_1}{r_1}
 \chainedlink{\overline{a}_i}{\bf  \overline{a}_i}\chainedlink{\bf \overline{a}_o}{\overline{a}_o}
 \chainedlink{r_2}{r_2}
 \chainedlink{q_i}{\bf q_i}\chainedlink{\bf q_o}{q_o}
 \chainedlink{b_i}{\bf b_i}\chainedlink{\bf b_o}{b_o}
 \chainedlink{\overline{w}_i}{\bf \overline{w}_i}\chainedlink{\bf \overline{w}_o}{\overline{w}_o}
 \chainedlink{\overline{a}_i}{\bf \overline{a}_i}\chainedlink{\bf \overline{a}_o}{\overline{a}_o}
 \chainedlink{r_3}{r_3}
 \chainedlink{\overline{a}_i}{\bf \overline{a}_i}\chainedlink{\bf \overline{a}_o}{\overline{a}_o}
 \chainedlink{r_4}{r_4}
 \chainedlink{\overline{w}_i}{\bf \overline{w}_i}\chainedlink{\bf \overline{w}_o}{\overline{w}_o}
 \chainedlink{\mathit{cxt}}{\mathit{cxt}}
 \chainedlink{\widehat{a}_i}{\bf \widehat{a}_i}\chainedlink{\bf \widehat{a}_o}{\widehat{a}_o}
 \chainedlink{\underline{b}_i}{\bf \underline{b}_i}\chainedlink{\bf \underline{b}_o}{\underline{b}_o}
 \chainedlink{p_1}{p_1}
 \chainedlink{p_2}{p_2}
 \chainedlink{\widetilde{q}_i}{\bf \widetilde{q}_i}\chainedlink{\bf \widetilde{q}_o}{\widetilde{q}_o}
 \chainedlink{p_3}{p_3}
 \chainedlink{p_4}{p_4}
 \chainend{\silent}
%\end{array}
\]}
The parts in bold are provided by the entity processes, the other parts are provided by the processes encoding the reactions and by the process encoding the context (starting at $\mathit{cxt}$ and ending at $p_1$.
In the label we can read that reactions $1$ and $4$ have been not executed because the entity $a$ is absent, 
the reaction $3$ has been not applied because the entity $w$ is absent, then only reaction $2$ has been applied, and it has produced the entity $q$. Also,  the context provides entity $a$, that will be available in the next state, and not the entity $b$.
Now, to let the label more readable, we show the result of the application of the function $flat(\cdot)$ to it:
\[
r_1~\overline{a}~r_2~q~b~\overline{w}~\overline{a}~r_3~\overline{a}~r_4~\overline{w}~\mathit{cxt}~\widehat{a}~\underline{b}~p_1~p_2~\widetilde{q}~p_3~p_4 .
\]

>>>>

Then, only the second reaction can be applied, and the transition carries the complete label $\upsilon$ below
{\tiny
\[
%\begin{array}{l}
%\upsilon = \\
%\restrict{r1,r2,r3,r4,\widetilde{q},\widetilde{w},\widetilde{a},\widetilde{b}}\\
 \startchain{\silent}
 \chainedlink{r_1}{r_1}
 \chainedlink{\overline{a}_i}{\bf  \overline{a}_i}\chainedlink{\bf \overline{a}_o}{\overline{a}_o}
 \chainedlink{r_2}{r_2}
 \chainedlink{q_i}{\bf q_i}\chainedlink{\bf q_o}{q_o}
 \chainedlink{b_i}{\bf b_i}\chainedlink{\bf b_o}{b_o}
 \chainedlink{\overline{w}_i}{\bf \overline{w}_i}\chainedlink{\bf \overline{w}_o}{\overline{w}_o}
 \chainedlink{\overline{a}_i}{\bf \overline{a}_i}\chainedlink{\bf \overline{a}_o}{\overline{a}_o}
 \chainedlink{r_3}{r_3}
 \chainedlink{\overline{a}_i}{\bf \overline{a}_i}\chainedlink{\bf \overline{a}_o}{\overline{a}_o}
 \chainedlink{r_4}{r_4}
 \chainedlink{\overline{w}_i}{\bf \overline{w}_i}\chainedlink{\bf \overline{w}_o}{\overline{w}_o}
 \chainedlink{\mathit{cxt}}{\mathit{cxt}}
 \chainedlink{\widehat{a}_i}{\bf \widehat{a}_i}\chainedlink{\bf \widehat{a}_o}{\widehat{a}_o}
 \chainedlink{\underline{b}_i}{\bf \underline{b}_i}\chainedlink{\bf \underline{b}_o}{\underline{b}_o}
 \chainedlink{p_1}{p_1}
 \chainedlink{p_2}{p_2}
 \chainedlink{\widetilde{q}_i}{\bf \widetilde{q}_i}\chainedlink{\bf \widetilde{q}_o}{\widetilde{q}_o}
 \chainedlink{p_3}{p_3}
 \chainedlink{p_4}{p_4}
 \chainend{\silent}
%\end{array}
\]}
The parts in bold are provided by the entity processes, the other parts are provided by the processes encoding the reactions and by the process encoding the context (starting at $\mathit{cxt}$ and ending at $p_1$).
In the label we can read that reactions $1$ and $4$ have been not executed because the entity $a$ is absent, 
the reaction $3$ has been not applied because the entity $w$ is absent, then only reaction $2$ has been applied, and it has produced the entity $q$. Also,  the context provides entity $a$, that will be available in the next state, and not the entity $b$.
Now, to let the label more readable, we show the result of the application of the function $flat(\cdot)$ to it:
\[
r_1~\overline{a}~r_2~q~b~\overline{w}~\overline{a}~r_3~\overline{a}~r_4~\overline{w}~\mathit{cxt}~\widehat{a}~\underline{b}~p_1~p_2~\widetilde{q}~p_3~p_4 .
\]
